---
title: "Assignment 1 - Dataset Selection"
author: "Aron Lesser"
date: "9/6/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    
  
---

```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
```

# Dataset Selection

After downloading the *tidycensus* and *tidyverse* libraries, I searched through person- and household-level data to define a set of variables. Using the include=FLASE command, they will run with my program but not appear on the knit file. 


### Loading Data


#### Person-level data

```{r}
person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs1")
```


#### Household-level data 

```{r}
hh_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "housing", year == 2018, survey == "acs1")
```
 
## Defining Variables

I selected a dataset that may allow me to examine mobility status and moving trends among renters of hispanic origin in Georgia. In this dataset, I use seven variables. 

In order to only include renters, I filtered for rent payments greater than zero. By doing so, I eliminated the housing tenure category altogether, which was in a previous draft.  

*Categorical*

1. **HISP**: Recoded detailed Hispanic origin
2. **MIG**: Mobility status (lived here 1 year ago)
3. **LANP**: Language spoken at home

*Continuous*

4. **PINCP**: Income
5. **NP**: Number of persons in this household
6. **AGEP**: Age
7. **RNTP**: Monthly rent 

## Downloading the data

I downloaded the data using the following code: 

```{r, message=FALSE, results='hide'}
person_data <- get_pums(variables = c("HISP", 
                                      "MIG", 
                                      "LANP", 
                                      "PINCP",
                                      "NP",
                                      "AGEP",
                                      "RNTP"),
                        state = "GA", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(RNTP > 0) %>%
  select(HISP_label, MIG_label, LANP, PINCP, NP, RNTP, AGEP)
```


# Finalizing and Writing

### First 10 Observations

Below are the first 10 observations of the dataset
```{r} 
person_data
```


### Writing a CSV File

This is the code that I used to write it into a .csv file: 

```{r}
write_csv(person_data, "people.csv")
```
