---
title: "Quant assignment 2"
author: "Aron Lesser"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggthemes)
```

Test change

# Loading Data

I begin by loading the data, filtering income to be greater than 0. 

```{r}
person_data <- read_csv("people.csv")%>%
  filter(PINCP > 0)
attach(person_data)
```

# Variables

For this assignment, I will use 5 variables:  

*Categorical*

1. **HISP**: Recoded detailed Hispanic origin
2. **MIG**: Mobility status (lived here 1 year ago)

*Continuous*

3. **PINCP**: Income
4. **NP**: Number of persons in this household
5. **RNTP**: Monthly rent 

I decided to exclude primary language spoken at home and age because I did not feel that they will add to my analysis moving forward. 

## Continuous Variables 

### Monthly Rent, Income, and Number of Persons in the Household

Using the summary function, I calculate the Median, Mean, and the 1st and 3rd quartiles. 
```{r}
summary(RNTP)

summary(PINCP)

summary(NP)
```

I then calculate the standard deviation for each variable
```{r}
sd(RNTP)
sd(PINCP)
sd(NP)
```

And run a T test to get the 95% confidence interval of each

T test Rent

```{r}
pop_mean <- t.test(RNTP)

pop_mean
```


T test Income

```{r}
pop_mean <- t.test(PINCP)

pop_mean
```

T test Number of Persons in Household


```{r}
pop_mean <- t.test(NP)

pop_mean
```


### Representing the Data on Histograms

Finally, I create histograms representing each continuous variables on the X axis and the number of people in the sample on the Y axis. I export it to pdf in order to use it in my submission InDesign Submission. 

#### Rent

```{r}
pdf("Rent_People_Sample.pdf", width = 3, height = 3) 

hist <- ggplot(person_data, aes(x = RNTP)) +
  geom_histogram() 

hist + 
  theme_tufte() +
  scale_x_continuous(name = "Amount spent on rent", limits = c(0, 3000)) +
  scale_y_continuous(name = "Number of people in sample")
dev.off()
```

#### Income
```{r}
pdf("income_People_Sample.pdf", width = 3.5, height = 3) 
hist <- ggplot(person_data, aes(x = PINCP)) +
  geom_histogram(bins = 20) 

hist + 
  theme_tufte() +
scale_x_continuous((name = "Income"), label = comma, limits = c(-0, 300000))+
  scale_y_continuous(name = "Number of people in sample")
dev.off()


```


#### Number of People in Household

```{r}
pdf("NP_household_histogram.pdf", width = 3, height = 3) 
   

hist <- ggplot(person_data, aes(x = NP)) +
  geom_histogram(bins=12) 

hist + 
  theme_tufte() +
  scale_x_continuous(name = "Number of people in each household", limits = c(0,12), breaks = c(1,3,5,7,9,11)) +
  scale_y_continuous(name = "Number of people in sample")

dev.off()


```





## Categorical Variables

### Mobility

I begin by displaying the categories for Mobility 

```{r}
mobility <- unique(MIG_label)

mobility
```

The I get the share of the sample in each category
```{r}
different_us <- t.test(MIG_label == mobility[1])
nonmovers <- t.test(MIG_label == mobility[2])
different_international <- t.test(MIG_label == mobility[3])



shares <- tibble(mobility = c("different_us", 
                            "nonmovers", 
                            "different_international"),
                 `Share` = c(different_us$estimate, 
                            nonmovers$estimate, 
                            different_international$estimate),
                 Low = c(different_us$conf.int[1], 
                            nonmovers$conf.int[1], 
                            different_international$conf.int[1]),
                 High = c(different_us$conf.int[2], 
                            nonmovers$conf.int[2], 
                            different_international$conf.int[2]))
shares
```
I also illustrated the results on a bar chart, just to familiarize myself with this type of visualization

```{r}
ggplot(shares, aes(x = mobility, y = Share, ymin = Low, ymax = High)) +
  geom_bar(stat = "identity") +
  geom_errorbar(width = 0.5) +
  theme_bw() +
  scale_y_continuous(name = "Share of population (error bars represent\n 95-percent confidence interval)",
                     breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8),
                     labels = c("0", "10%", "20%", "30%", "40%", "50%", "60%", "70", "80"))
```

### Ethnicity of Hispanic Origin

I repet the process to look at the label categories for Hispanic Origin

```{r}
h_origin <- unique(HISP_label)

h_origin
```

I then get the share of the sample in each category

```{r}
Not_Spanish_Hispanic_Latino <- t.test(HISP_label == h_origin[1])
Dominican <- t.test(HISP_label == h_origin[2])
Guatemalan <- t.test(HISP_label == h_origin[3])
Mexican <-  t.test(HISP_label == h_origin[4])
Salvadoran <-  t.test(HISP_label == h_origin[5])
Ecuadorian <-  t.test(HISP_label == h_origin[6])
Puerto_Rican <-  t.test(HISP_label == h_origin[7])
All_other_latino <-  t.test(HISP_label == h_origin[8])
Colombian <-  t.test(HISP_label == h_origin[9])
Panamanian <-  t.test(HISP_label == h_origin[10])
Cuban <-  t.test(HISP_label == h_origin[11])
Honduran <-  t.test(HISP_label == h_origin[12])
Peruvian <-  t.test(HISP_label == h_origin[13])
Spaniard <-  t.test(HISP_label == h_origin[14])
Venezuelan <-  t.test(HISP_label == h_origin[15])
Uruguayan <-  t.test(HISP_label == h_origin[16])
Nicaraguan <-  t.test(HISP_label == h_origin[17])
Costa_Rican <-  t.test(HISP_label == h_origin[18])
Other_Central_American <-  t.test(HISP_label == h_origin[19])
Argentinian <-  t.test(HISP_label == h_origin[20])
Chilean <-  t.test(HISP_label == h_origin[21])
Paraguayan <-  t.test(HISP_label == h_origin[22])
Other_South_America <-  t.test(HISP_label == h_origin[23])


shares <- tibble(h_origin = c("Not_Spanish_Hispanic_Latino",
                              "Dominican",
                              "Guatemalan",
                              "Mexican",
                              "Salvadoran",
                              "Ecuadorian",
                              "Puerto_Rican", 
                              "All_other_latino", 
                              "Colombian",
                              "Panamanian",
                              "Cuban",
                              "Honduran",
                              "Peruvian",
                              "Spaniard",
                              "Venezuelan",
                              "Uruguayan",
                              "Nicaraguan", 
                              "Costa_Rican",
                              "Other_Central_American",
                              "Argentinian",
                              "Chilean",
                              "Paraguayan", 
                              "Other_South_America"),
                 
                 `Share` = c(Not_Spanish_Hispanic_Latino$estimate,
                               Dominican$estimate,
                               Guatemalan$estimate, 
                               Mexican$estimate,
                               Salvadoran$estimate, 
                               Ecuadorian$estimate, 
                               Puerto_Rican$estimate, 
                               All_other_latino$estimate, 
                               Colombian$estimate, 
                               Panamanian$estimate, 
                               Cuban$estimate, 
                               Honduran$estimate,
                               Peruvian$estimate, 
                               Spaniard$estimate, 
                               Venezuelan$estimate, 
                               Uruguayan$estimate, 
                               Nicaraguan$estimate, 
                               Costa_Rican$estimate, 
                               Other_Central_American$estimate, 
                               Argentinian$estimate, 
                               Chilean$estimate,
                               Paraguayan$estimate,
                               Other_South_America$estimate),
                 
                 Low = c(Not_Spanish_Hispanic_Latino$conf.int[1],
                               Dominican$conf.int[1],
                               Guatemalan$conf.int[1],
                               Mexican$conf.int[1],
                               Salvadoran$conf.int[1],
                               Ecuadorian$conf.int[1],
                               Puerto_Rican$conf.int[1],
                               All_other_latino$conf.int[1],
                               Colombian$conf.int[1],
                               Panamanian$conf.int[1],
                               Cuban$conf.int[1],
                               Honduran$conf.int[1],
                               Peruvian$conf.int[1],
                               Spaniard$conf.int[1],
                               Venezuelan$conf.int[1],
                               Uruguayan$conf.int[1],
                               Nicaraguan$conf.int[1],
                               Costa_Rican$conf.int[1],
                               Other_Central_American$conf.int[1],
                               Argentinian$conf.int[1],
                               Chilean$conf.int[1],
                               Paraguayan$conf.int[1],
                               Other_South_America$conf.int[1]),
                 
                 High = c(Not_Spanish_Hispanic_Latino$conf.int[2],
                               Dominican$conf.int[2],
                               Guatemalan$conf.int[2],
                               Mexican$conf.int[2],
                               Salvadoran$conf.int[2],
                               Ecuadorian$conf.int[2],
                               Puerto_Rican$conf.int[2],
                               All_other_latino$conf.int[2],
                               Colombian$conf.int[2],
                               Panamanian$conf.int[2],
                               Cuban$conf.int[2],
                               Honduran$conf.int[2],
                               Peruvian$conf.int[2],
                               Spaniard$conf.int[2],
                               Venezuelan$conf.int[2],
                               Uruguayan$conf.int[2],
                               Nicaraguan$conf.int[2],
                               Costa_Rican$conf.int[2],
                               Other_Central_American$conf.int[2],
                               Argentinian$conf.int[2],
                               Chilean$conf.int[2],
                               Paraguayan$conf.int[2],
                               Other_South_America$conf.int[2]))
shares
```


# Finalizing the Submisison

To finalize the submission, I transfer the data to my InDesign layout and link the histograms. I include my discussion questions and reflections on the data in the final pdf submission. 

