---
title: "Quant Assignment 3"
author: "Aron Lesser"
date: "9/22/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggthemes)
```


# Loading Data

I begin by loading the data and filtering income to be greater than 0. I also altered the ethnicity category to only include Hispanic/Not Hispanic.

```{r}
new_person_data <- read_csv("people.csv")%>%
   filter(PINCP > 0)%>%
  mutate(HISP_label = case_when(
    HISP_label == "Not Spanish/Hispanic/Latino" ~ "Non Hispanic",
    HISP_label ==                          "Dominican" ~ "Hispanic",
    HISP_label ==                          "Guatemalan" ~ "Hispanic",
    HISP_label ==                          "Mexican" ~ "Hispanic",
    HISP_label ==                          "Salvadoran" ~ "Hispanic",
    HISP_label ==                          "Ecuadorian" ~ "Hispanic",
    HISP_label ==                          "Puerto Rican" ~ "Hispanic", 
    HISP_label ==                          "All Other Spanish/Hispanic/Latino" ~ "Hispanic", 
    HISP_label ==                          "Colombian" ~ "Hispanic",
    HISP_label ==                          "Panamanian" ~ "Hispanic",
    HISP_label ==                          "Cuban" ~ "Hispanic",
    HISP_label ==                          "Honduran" ~ "Hispanic",
    HISP_label ==                          "Peruvian" ~ "Hispanic",
    HISP_label ==                          "Spaniard" ~ "Hispanic",
    HISP_label ==                          "Venezuelan" ~ "Hispanic",
    HISP_label ==                          "Uruguayan" ~ "Hispanic",
    HISP_label ==                          "Nicaraguan" ~ "Hispanic", 
    HISP_label ==                          "Costa Rican" ~ "Hispanic",
    HISP_label ==                          "Other Central American" ~ "Hispanic",
    HISP_label ==                          "Argentinean" ~ "Hispanic",
    HISP_label ==                          "Chilean" ~ "Hispanic",
    HISP_label ==                          "Paraguayan" ~ "Hispanic", 
    HISP_label ==                          "Other South American" ~ "Hispanic"))
  attach(new_person_data)
```

# Variables

For this assignment, I will use 5 variables:  

*Categorical*

1. **HISP**: Altered to include only Hispanic and Non-Hispanic
2. **MIG**: Mobility status (lived here 1 year ago)


*Continuous*

3. **PINCP**: Income
4. **NP**: Number of persons in this household
5. **RNTP**: Monthly rent 


# Relationships Between Two Continuous Variables

## Correlation between Income and Rent  
```{r}
correlation <- cor.test(new_person_data$PINCP, new_person_data$RNTP)

correlation
```

I was expecting to see a stronger correlation betweeen income and rent, which is only .3 

### Plotting Income and Rent 

```{r}
ggplot(new_person_data, aes(x = PINCP, y = RNTP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_minimal() +
  scale_x_continuous(name = "Income", label = comma) +
  scale_y_continuous(name = "Monthly rent", 
                     breaks = seq(0, 4000, by = 500),
                     labels = paste("$", 
                                    prettyNum(seq(0, 4000, by = 500),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 0, y = 4000, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```

### Testing Income and Rent at a max income  of $100,000

It appeared that there may be a slightly different correlation strength among people in the sample that earn $100,000 or less. I created a new dataset to test this hypothesis. 

```{r}
new_person_data_1 <- read_csv("people.csv") %>%
   filter(PINCP > 0, PINCP < 100000) %>%
  mutate(HISP_label = case_when(
    HISP_label == "Not Spanish/Hispanic/Latino" ~ "Non Hispanic",
    HISP_label ==                          "Dominican" ~ "Hispanic",
    HISP_label ==                          "Guatemalan" ~ "Hispanic",
    HISP_label ==                          "Mexican" ~ "Hispanic",
    HISP_label ==                          "Salvadoran" ~ "Hispanic",
    HISP_label ==                          "Ecuadorian" ~ "Hispanic",
    HISP_label ==                          "Puerto Rican" ~ "Hispanic", 
    HISP_label ==                          "All Other Spanish/Hispanic/Latino" ~ "Hispanic", 
    HISP_label ==                          "Colombian" ~ "Hispanic",
    HISP_label ==                          "Panamanian" ~ "Hispanic",
    HISP_label ==                          "Cuban" ~ "Hispanic",
    HISP_label ==                          "Honduran" ~ "Hispanic",
    HISP_label ==                          "Peruvian" ~ "Hispanic",
    HISP_label ==                          "Spaniard" ~ "Hispanic",
    HISP_label ==                          "Venezuelan" ~ "Hispanic",
    HISP_label ==                          "Uruguayan" ~ "Hispanic",
    HISP_label ==                          "Nicaraguan" ~ "Hispanic", 
    HISP_label ==                          "Costa Rican" ~ "Hispanic",
    HISP_label ==                          "Other Central American" ~ "Hispanic",
    HISP_label ==                          "Argentinean" ~ "Hispanic",
    HISP_label ==                          "Chilean" ~ "Hispanic",
    HISP_label ==                          "Paraguayan" ~ "Hispanic", 
    HISP_label ==                          "Other South American" ~ "Hispanic"))
  attach(new_person_data_1)
```

### Correlation between 100k income and rent

As it turns out, it is not significantly different than above. I will need to think about some possible explanations. It is clear that there are more renters in lower income brackets. 

```{r}
correlation <- cor.test(new_person_data_1$PINCP, new_person_data_1$RNTP)

correlation
```


### Plotting 100k Income and Rent 


```{r}
ggplot(new_person_data_1, aes(x = PINCP, y = RNTP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_minimal() +
  scale_x_continuous(name = "Income", label = comma) +
  scale_y_continuous(name = "Monthly rent", 
                     breaks = seq(0, 4000, by = 500),
                     labels = paste("$", 
                                    prettyNum(seq(0, 4000, by = 500),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 0, y = 4000, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```



## Correlation between Household Size and Rent  
```{r}
correlation <- cor.test(new_person_data$NP, new_person_data$RNTP)

correlation
```
There is an extremely modest correlation between household size and monthly rent. 

### Plotting Household Size and Rent 

```{r}
ggplot(new_person_data, aes(x = NP, y = RNTP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Household Size", limits = c(1,12), seq(1, 12, by = 2)) +
  scale_y_continuous(name = "Monthly rent", 
                     breaks = seq(0, 4000, by = 500),
                     labels = paste("$", 
                                    prettyNum(seq(0, 4000, by = 500),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1, y = 3500, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```

## Correlation between Household Size and Income  
```{r}
correlation <- cor.test(new_person_data$NP, new_person_data$PINCP)

correlation
```

This is again an interesting case in which there seems to be a very weak correlation between household size and income. 

### Plotting Household Size and Income 

```{r}

options(scipen=999)

ggplot(new_person_data, aes(x = NP, y = PINCP)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Household Size", limits = c(1,12), seq(1, 12, by = 2)) +
  scale_y_continuous(name = "Income",
                     breaks = seq(0, 600000, by = 50000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 600000, by = 50000),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1, y = 600000, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation$conf.int[2], digits = 2)),
           hjust = 0)
```

# Relationships between a Continuous Variable and a Binary Categorical One - Ethnicity 

## Ethnicity and Income

```{r}
difference = t.test(PINCP ~ HISP_label == "Hispanic",
                    data = new_person_data)

difference
```

There appears to be a statistically significant relationship between ethnicity and income. In this case, the mean income among Hispanics is nearly 10% higher than that of non-Hispanics. 

### Plotting Ethnicity and Income

```{r}
options(scipen=999)

ggplot(new_person_data, aes(x = HISP_label, y = PINCP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Ethnicity",
                   labels = c("Non-Hispanic", "Hispanic")) +
   scale_y_continuous(name = "Income", 
                     breaks = seq(0, 600000, by = 50000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 600000, by = 50000),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1.5, y = 250000, size = 3,
           label = paste("Difference in means\n(95% confidence):\n$",
                         prettyNum(abs(difference$conf.int[2]), 
                                   digits = 0), " to $",
                         prettyNum(abs(difference$conf.int[1]), 
                                   digits = 0),
                         sep = "")) +
  annotate(geom = "text", x = 1, y = 26000, size = 2,
           label = paste("Mean (95% confidence):\n$",
                         prettyNum(difference$estimate[1], digits = 0))) +
  annotate(geom = "text", x = 2, y = 26000, size = 2,
           label = paste("Mean (95% confidence):\n$",
                         prettyNum(difference$estimate[2], digits = 0)))
```


## Ethnicity and Rent

```{r}
difference = t.test(RNTP ~ HISP_label == "Hispanic",
                    data = new_person_data)

difference
```

Interestingly, while Non-Hispanics in this sample on average make more than Hispanics, Non-Hispanics pay on average ~10% less rent than Hispanics. The t-test suggests that this is a statistically significant relationship.  

### Plotting Ethnicity and Rent

```{r}
ggplot(new_person_data, aes(x = HISP_label, y = RNTP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Ethnicity",
                   labels = c("Non-Hispanic", "Hispanic")) +
   scale_y_continuous(name = "Rent", 
                     breaks = seq(0, 4000, by = 500),
                     labels = paste("$", 
                                    prettyNum(seq(0, 4000, by = 500),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1.5, y = 2500, size = 3,
           label = paste("Difference in means\n(95% confidence):\n$",
                         prettyNum(abs(difference$conf.int[2]), 
                                   digits = 0), " to $",
                         prettyNum(abs(difference$conf.int[1]), 
                                   digits = 0),
                         sep = "")) +
  annotate(geom = "text", x = 1, y = difference$estimate[1], size = 3,
           label = paste("Mean (95% confidence):\n$",
                         prettyNum(difference$estimate[1], digits = 0))) +
  annotate(geom = "text", x = 2, y = difference$estimate[2], size = 3,
           label = paste("Mean (95% confidence):\n$",
                         prettyNum(difference$estimate[2], digits = 0)))
```



## Ethnicity and Household Size

```{r}
difference = t.test(NP ~ HISP_label == "Hispanic",
                    data = new_person_data)

difference
```
It appears that there is a statistically significant relationship between ethnicity and household size. In this case, Hispanics households are on average 1 person greater than their non-Hispanic counterparts. 


### Plotting Ethnicity and Household Size

```{r}
ggplot(new_person_data, aes(x = HISP_label, y = NP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Ethnicity",
                   labels = c("Non-Hispanic", "Hispanic")) +
   scale_y_continuous(name = "Household Size", 
                     breaks = seq(1, 12, by = 1),
                     labels = paste(
                                    prettyNum(seq(1, 12, by = 1),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1.5, y = 8, size = 3,
           label = paste("Difference in means\n(95% confidence):\n",
                         prettyNum(abs(difference$conf.int[2]), 
                                   digits = 0),
                         prettyNum(abs(difference$conf.int[1]), 
                                   digits = 0),
                         sep = "")) +
  annotate(geom = "text", x = .6, y = difference$estimate[1], size = 2,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[1], digits = 0))) +
  annotate(geom = "text", x = 2.4, y = difference$estimate[2], size = 2,
           label = paste("Mean (95% confidence):\n",
                         prettyNum(difference$estimate[2], digits = 0)))
```


I realize that in plotting this on a graph, the difference in means comes out as 11. I do not understand why this is happening, to be honest, since the 95% confidence interval is -1 to -.08. Could it be the negatives? 


# Relationships between a Continuous Variable and a Non-Binary Categorical One - Mobility Status

## Anova Mobility Status and Rent

```{r}
anova <- aov(RNTP ~ MIG_label, data = new_person_data)

summary(anova)
```

There appears to be a statistically significant relationship between the variables. 

### Tukey's Honestly Significant Difference - Mobility Status and Rent

```{r}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$MIG_label),
                differences$MIG_label))
```

### Plotting Mobility Status and Rent

In this case, I found a boxplot to be more effective than a violin plot. 

```{r}
ggplot(new_person_data, aes(x = MIG_label, y = RNTP)) +
  geom_boxplot() +
  theme_bw() +
  scale_x_discrete(name = "Mobility Status",
   labels = c("Moved within U.S.", "Moved Internationally", "Did Not Move")) +
  scale_y_continuous(name = "Rent", 
                     breaks = seq(0, 4000, by = 500),
                     labels = paste("$", 
                                    prettyNum(seq(0, 4000, by = 500),
                                              big.mark = ","))) 
```


In this boxplot we can see differences in medians and interquartile ranges. 

## Anova Mobility Status and Income

```{r}
anova <- aov(PINCP ~ MIG_label, data = new_person_data)

summary(anova)
```


While there is a statistically significant relationship, the p value is much lower among these variables than the previous ones. 

### Tukey's Honestly Significant Difference - Mobility Status and Income

```{r}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$MIG_label),
                differences$MIG_label))
```

### Plotting Mobility Status and Income


```{r}
ggplot(new_person_data, aes(x = MIG_label, y = PINCP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Mobility Status",
   labels = c("Moved within U.S.", "Moved Internationally", "Did Not Move")) +
  scale_y_continuous(name = "Income", 
                     breaks = seq(0, 600000, by = 50000),
                     labels = paste("$", 
                                    prettyNum(seq(0, 600000, by = 50000),
                                              big.mark = ","))) 
```

## Anova Mobility Status and Household Size

```{r}
anova <- aov(NP ~ MIG_label, data = new_person_data)

summary(anova)
```

### Tukey's Honestly Significant Difference - Mobility Status and Household Size

```{r}
differences <- TukeyHSD(anova)

as_tibble(cbind(pair = row.names(differences$MIG_label),
                differences$MIG_label))
```

### Plotting Mobility Status and Household Size


```{r}
ggplot(new_person_data, aes(x = MIG_label, y = NP)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Mobility Status",
   labels = c("Moved within U.S.", "Moved Internationally", "Did Not Move")) +
  scale_y_continuous(name = "Household Size", 
                     breaks = seq(1, 12, by = 1),
                     labels = paste(
                                    prettyNum(seq(1, 12, by = 1),
                                              big.mark = ",")))
```


# Two Categorial Variables - Mobility Status and Ethnicity

## Hispanic Ethnicity and Mobility Status

```{r}
chi_sq <- chisq.test(new_person_data$MIG_label, new_person_data$HISP_label == 
                       "Hispanic") 

chi_sq
```

In running a Chi-squared test, it appears that there is a statistically significant relationship between ethnicity and mobility status, both for Hispanics and non-Hispanics. However, I would like to better understand how to interpret the results. 

Observed

```{r}
chi_sq$observed
```

Expected

```{r}
chi_sq$expected
```

## Non-Hispanic Ethnicity and Mobility Status

```{r}
chi_sq <- chisq.test(new_person_data$MIG_label, new_person_data$HISP_label == 
                       "Non Hispanic") 

chi_sq
```

There also appears to be a statistically significant relationship between the variables of non-Hispanic Ethnicity and mobility status. 

Observed

```{r}
chi_sq$observed
```

Expected

```{r}
chi_sq$expected
```