---
title: "Quant Assignment 4"
author: "Aron Lesser"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggthemes)
```


# Loading Data

I begin by loading the data and filtering income to be greater than 0. I also altered the ethnicity category to only include Hispanic/Not Hispanic.

```{r}
new_person_data_regressions <- read_csv("people.csv")%>%
   filter(PINCP > 0)%>%
  mutate(HISP_label = case_when(
    HISP_label == "Not Spanish/Hispanic/Latino" ~ "Non Hispanic",
    HISP_label ==                          "Dominican" ~ "Hispanic",
    HISP_label ==                          "Guatemalan" ~ "Hispanic",
    HISP_label ==                          "Mexican" ~ "Hispanic",
    HISP_label ==                          "Salvadoran" ~ "Hispanic",
    HISP_label ==                          "Ecuadorian" ~ "Hispanic",
    HISP_label ==                          "Puerto Rican" ~ "Hispanic", 
    HISP_label ==                          "All Other Spanish/Hispanic/Latino" ~ "Hispanic", 
    HISP_label ==                          "Colombian" ~ "Hispanic",
    HISP_label ==                          "Panamanian" ~ "Hispanic",
    HISP_label ==                          "Cuban" ~ "Hispanic",
    HISP_label ==                          "Honduran" ~ "Hispanic",
    HISP_label ==                          "Peruvian" ~ "Hispanic",
    HISP_label ==                          "Spaniard" ~ "Hispanic",
    HISP_label ==                          "Venezuelan" ~ "Hispanic",
    HISP_label ==                          "Uruguayan" ~ "Hispanic",
    HISP_label ==                          "Nicaraguan" ~ "Hispanic", 
    HISP_label ==                          "Costa Rican" ~ "Hispanic",
    HISP_label ==                          "Other Central American" ~ "Hispanic",
    HISP_label ==                          "Argentinean" ~ "Hispanic",
    HISP_label ==                          "Chilean" ~ "Hispanic",
    HISP_label ==                          "Paraguayan" ~ "Hispanic", 
    HISP_label ==                          "Other South American" ~ "Hispanic")) %>%
 
 mutate(MIG_label = case_when(
    MIG_label == "Yes, same house (nonmovers)" ~ "1Yes, same house (nonmovers)",
    MIG_label == "No, different house in US or Puerto Rico" ~ "No, different house in US or Puerto Rico",
    MIG_label == "No, outside US and Puerto Rico" ~ "No, outside US and Puerto Rico"))
  attach(new_person_data_regressions)
```
# Variables

For this assignment, I will use 5 variables:  

*Categorical*

1. **HISP**: Altered to include only Hispanic and Non-Hispanic
2. **MIG**: Mobility status (lived here 1 year ago)


*Continuous*

3. **PINCP**: Income
4. **NP**: Number of persons in this household
5. **RNTP**: Monthly rent 

```{r}
model <- lm(PINCP ~ NP + RNTP + MIG_label + HISP_label, 
            data = new_person_data_regressions)

summary(model)
```

# Interpreting the results

I ran a regression to predict the relationship between income and my other variables. With an R squared of 1.074, it appears that for every additional person in the household, renters will make nearly $3,000 less. 

It also predicts that for every additional dollar spent on rent, renters make $26 more for their income.

Regarding migratory status, renters that move within the us earn ~2,000 less than those that stay in their homes. Renters that move outside of the U.S. earn ~6,700 less than those that stay in their homes. Here we see a realtionship between higher income and not moving. 

Finally, it appears that non-hispanic renters will earn $2,845 more than Hispanics? It appeared as though the mean salary of Hispanic renters was higher than that of non-hispanic renters, so I need to review this finding. 
