---
title: "Quant Assignment 4"
author: "Aron Lesser"
date: "9/30/2020"
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
library(ggplot2)
library(scales)
library(ggthemes)
```


# Loading Data

I begin by loading the data and filtering income to be greater than 0. I also altered the ethnicity category to only include Hispanic/Not Hispanic. I also altered the Migratory Status options to make "did not move" as the base for the analysis. 

```{r}
new_person_data_regressions <- read_csv("people.csv")%>%
   filter(PINCP > 0)%>%
  mutate(HISP_label = case_when(
    HISP_label == "Not Spanish/Hispanic/Latino" ~ "Non Hispanic",
    HISP_label ==                          "Dominican" ~ "Hispanic",
    HISP_label ==                          "Guatemalan" ~ "Hispanic",
    HISP_label ==                          "Mexican" ~ "Hispanic",
    HISP_label ==                          "Salvadoran" ~ "Hispanic",
    HISP_label ==                          "Ecuadorian" ~ "Hispanic",
    HISP_label ==                          "Puerto Rican" ~ "Hispanic", 
    HISP_label ==                          "All Other Spanish/Hispanic/Latino" ~ "Hispanic", 
    HISP_label ==                          "Colombian" ~ "Hispanic",
    HISP_label ==                          "Panamanian" ~ "Hispanic",
    HISP_label ==                          "Cuban" ~ "Hispanic",
    HISP_label ==                          "Honduran" ~ "Hispanic",
    HISP_label ==                          "Peruvian" ~ "Hispanic",
    HISP_label ==                          "Spaniard" ~ "Hispanic",
    HISP_label ==                          "Venezuelan" ~ "Hispanic",
    HISP_label ==                          "Uruguayan" ~ "Hispanic",
    HISP_label ==                          "Nicaraguan" ~ "Hispanic", 
    HISP_label ==                          "Costa Rican" ~ "Hispanic",
    HISP_label ==                          "Other Central American" ~ "Hispanic",
    HISP_label ==                          "Argentinean" ~ "Hispanic",
    HISP_label ==                          "Chilean" ~ "Hispanic",
    HISP_label ==                          "Paraguayan" ~ "Hispanic", 
    HISP_label ==                          "Other South American" ~ "Hispanic")) %>%
 
 mutate(MIG_label = case_when(
    MIG_label == "Yes, same house (nonmovers)" ~ "1Yes, same house (nonmovers)",
    MIG_label == "No, different house in US or Puerto Rico" ~ "No, different house in US or Puerto Rico",
    MIG_label == "No, outside US and Puerto Rico" ~ "No, outside US and Puerto Rico"))
  attach(new_person_data_regressions)
```
# Variables

For this assignment, I will use 5 variables:  

*Categorical*

1. **HISP**: Altered to include only Hispanic and Non-Hispanic
2. **MIG**: Mobility status (lived here 1 year ago)


*Continuous*

3. **PINCP**: Income
4. **NP**: Number of persons in this household
5. **RNTP**: Monthly rent 


# Regression

I estimate a regression model to predict the value of income based on the other variables. 


```{r}
model <- lm(PINCP ~ NP + RNTP + MIG_label + HISP_label, 
            data = new_person_data_regressions)

summary(model)
```

# Interpreting the results

With an R squared of 1.074 the model does predict outcomes with some degree of significance, especially in the field of Urban Planning.

The model predicts that for for every additional person in the household, renters would make make nearly $3,000 less. This relationship has a very low p value, so we understand the relationship between the variables to have a statistical significance with more than 95% confidence.  

The model also suggests that for every additional dollar spent on rent, renters earn ~$26 more for their income. This relationship has a very low p value, so we can say this is the case with well more than 95% confidence.  

Regarding migratory status, renters that move within the United States are predicted to earn ~2,000 less than those that stay in their homes with above a 95% confidence (the p value is 0.00161). Renters that move outside of the U.S. would earn ~6,700 less than those that stay in their homes, though the P value is higher than U.S. movers, though still below the .05 that we expect for a 95% confidence. Here we see a relationship between higher income and not moving overall.  

Finally, it appears that Non-hispanic renters will earn $2,845 more than Hispanics renters. This is consistent with the difference in means that I found earlier (in the previous assignment). The p value of 0.00380 suggests that there is above a 95% confidence that the relationship between the variables is meaningful.  

